### 2024-12-25 mixed-models.qmd
20241016 - figure alt text
20241225 - convert to calculate_stats() from nflfastR v5.0

::: {#fig-simpsonsParadox}
![](images/simpsonsParadox.png)\

@@

```{r}
player_stats_seasonal_offense_subset <- player_stats_seasonal_offense %>% 
  filter(position_group %in% c("QB","RB","WR","TE"))

@@

player_stats_seasonal_kicking_subset <- player_stats_seasonal_kicking %>% 
  filter(position == "K")

player_stats_seasonal_offense_subset <- bind_rows(
  player_stats_seasonal_offense_subset,
  player_stats_seasonal_kicking_subset
)

@@

```{r}
player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
  filter(
    !is.na(player_idFactor),
    !is.na(fantasy_points),
    !is.na(positionFactor),
    !is.na(ageCentered20),
    !is.na(ageCentered20Quadratic),
    !is.na(years_of_experience))
```

@@

plot_scatterplotFantasyPointsByAgeQB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "QB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeQB,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

```{r}
cor.test(
  formula = ~ age + fantasy_points,
  data = player_stats_seasonal_offense_subset %>% filter(position == "QB")
)
```

@@

plot_scatterplotFantasyPointsByAgeFB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "FB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Fullbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeFB,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

plot_scatterplotFantasyPointsByAgeRB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "RB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Running Backs"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeRB,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

plot_scatterplotFantasyPointsByAgeWR <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "WR") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeWR,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

plot_scatterplotFantasyPointsByAgeTE <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "TE") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Tight Ends"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeTE,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

```{r}
cor.test(
  formula = ~ age + fantasy_points,
  data = player_stats_seasonal_offense_subset %>% filter(position == "TE")
)
```

@@

plot_rawFantasyPointsByAgeQB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "QB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeQB,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

plot_rawFantasyPointsByAgeFB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "FB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Fullbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeFB,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

plot_rawFantasyPointsByAgeRB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "RB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Running Backs"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeRB,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

plot_rawFantasyPointsByAgeWR <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "WR") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeWR,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

plot_rawFantasyPointsByAgeTE <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "TE") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Tight Ends"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeTE,
  tooltip = c("age","fantasy_points","text","label"))
```

@@

```{r}
pointsPerSeason_nullModel <- lm(
  fantasy_points ~ 1,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

@@

```{r}
pointsPerSeason_linearRegression <- lm(
  fantasy_points ~ positionFactor + ageCentered20 + positionFactor:ageCentered20,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

@@

```{r}
pointsPerSeason_quadraticRegression <- lm(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

@@

```{r}
pointsPerSeason_randomIntercepts <- lmerTest::lmer(
  fantasy_points ~ 1 + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

plot_individualFantasyPointsRandomIntercepts <- ggplot(
  data = player_stats_seasonal_offense_subsetCC %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints_randomIntercepts = round(fantasyPoints_randomIntercepts, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints_randomIntercepts,
    group = player_id)) +
  geom_line(
    aes(
      x = age,
      y = fantasyPoints_randomIntercepts,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    linewidth = 0.5,
    color = "black") +
  geom_line(
    mapping = aes(
      x = age,
      y = fantasyPoints_randomIntercepts
    ),
    data = pointsPerSeason_positionAge_newData %>% 
      mutate(
        age = round(age, 2),
        fantasy_points = round(fantasyPoints_randomIntercepts, 2)
        ),
    inherit.aes = FALSE,
    se = TRUE,
    color = "#3366FF",
    linewidth = 2
  ) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position: Random Intercepts Model",
    #color = "Position"
  ) +
  theme_classic()

@@

```{r}
pointsPerSeason_position <- lmerTest::lmer(
  fantasy_points ~ positionFactor + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeFixedLinearSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)
```

@@

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

pointsPerSeason_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
  data = player_stats_seasonal_offense_subset,
  nthreads = num_cores
)

@@

```{r}
pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  control = lmerControl(optimizer = "bobyqa")
)

@@

pointsPerSeasonDepth_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
  data = player_stats_seasonal_offense_subsetDepth,
  nthreads = num_cores
)

@@

```{r}
player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasy_points
player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasy_points < 0] <- 0
player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasy_points <= 0] <- 0.01

fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasy_points %>% 
    na.omit() %>% 
    as.vector()

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_subset,
  mapping = aes(
    x = fantasy_points)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  theme_classic()
```

>>>>

::: {#fig-simpsonsParadox}
![](images/simpsonsParadox.png){fig-alt="An Example of Simpson's Paradox. In this example, there is a positive association between `x` and `y` within each group (i.e., red or blue), but there is a negative association between `x` and `y` when the groups are combined. (Figure retrieved from <https://en.wikipedia.org/wiki/File:Simpson%27s_paradox_continuous.svg>)"}

@@

```{r}
player_stats_seasonal_offense_subset <- player_stats_seasonal %>% 
  dplyr::filter(position_group %in% c("QB","RB","WR","TE") | position %in% c("K"))

@@

@@

```{r}
player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
  filter(
    !is.na(player_idFactor),
    !is.na(fantasyPoints),
    !is.na(positionFactor),
    !is.na(ageCentered20),
    !is.na(ageCentered20Quadratic),
    !is.na(years_of_experience))
```

@@

plot_scatterplotFantasyPointsByAgeQB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "QB") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasyPoints),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeQB,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

```{r}
cor.test(
  formula = ~ age + fantasyPoints,
  data = player_stats_seasonal_offense_subset %>% filter(position == "QB")
)
```

@@

plot_scatterplotFantasyPointsByAgeFB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "FB") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasyPoints),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Fullbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeFB,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

plot_scatterplotFantasyPointsByAgeRB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "RB") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasyPoints),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Running Backs"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeRB,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

plot_scatterplotFantasyPointsByAgeWR <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "WR") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasyPoints),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeWR,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

plot_scatterplotFantasyPointsByAgeTE <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "TE") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasyPoints),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Tight Ends"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeTE,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

```{r}
cor.test(
  formula = ~ age + fantasyPoints,
  data = player_stats_seasonal_offense_subset %>% filter(position == "TE")
)
```

@@

plot_rawFantasyPointsByAgeQB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "QB") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeQB,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

plot_rawFantasyPointsByAgeFB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "FB") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Fullbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeFB,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

plot_rawFantasyPointsByAgeRB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "RB") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Running Backs"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeRB,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

plot_rawFantasyPointsByAgeWR <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "WR") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeWR,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

plot_rawFantasyPointsByAgeTE <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "TE") %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints = round(fantasyPoints, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Tight Ends"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeTE,
  tooltip = c("age","fantasyPoints","text","label"))
```

@@

```{r}
pointsPerSeason_nullModel <- lm(
  fantasyPoints ~ 1,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

@@

```{r}
pointsPerSeason_linearRegression <- lm(
  fantasyPoints ~ positionFactor + ageCentered20 + positionFactor:ageCentered20,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

@@

```{r}
pointsPerSeason_quadraticRegression <- lm(
  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

@@

```{r}
pointsPerSeason_randomIntercepts <- lmerTest::lmer(
  fantasyPoints ~ 1 + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

plot_individualFantasyPointsRandomIntercepts <- ggplot(
  data = player_stats_seasonal_offense_subsetCC %>% 
    mutate(
      age = round(age, 2),
      fantasyPoints_randomIntercepts = round(fantasyPoints_randomIntercepts, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasyPoints_randomIntercepts,
    group = player_id)) +
  geom_line(
    aes(
      x = age,
      y = fantasyPoints_randomIntercepts,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    linewidth = 0.5,
    color = "black") +
  geom_line(
    mapping = aes(
      x = age,
      y = fantasyPoints_randomIntercepts
    ),
    data = pointsPerSeason_positionAge_newData %>% 
      mutate(
        age = round(age, 2),
        fantasyPoints = round(fantasyPoints_randomIntercepts, 2)
        ),
    inherit.aes = FALSE,
    se = TRUE,
    color = "#3366FF",
    linewidth = 2
  ) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position: Random Intercepts Model",
    #color = "Position"
  ) +
  theme_classic()

@@

```{r}
pointsPerSeason_position <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeFixedLinearSlopes <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + ageCentered20 + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)
```

@@

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

@@

pointsPerSeason_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
  fantasyPoints ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
  data = player_stats_seasonal_offense_subset,
  nthreads = num_cores
)

@@

```{r}
pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  control = lmerControl(optimizer = "bobyqa")
)

@@

pointsPerSeasonDepth_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
  fantasyPoints ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
  data = player_stats_seasonal_offense_subsetDepth,
  nthreads = num_cores
)

@@

```{r}
player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasyPoints
player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasyPoints < 0] <- 0
player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasyPoints <= 0] <- 0.01

fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasyPoints %>% 
    na.omit() %>% 
    as.vector()

@@

ggplot2::ggplot(
  data = player_stats_seasonal_offense_subset,
  mapping = aes(
    x = fantasyPoints)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  theme_classic()
```
